### 2020-01-13 script/data-prep-A.Rmd

---
title: 'Data preparation A'
author: ""
output:
    html_document:
        smart: false
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
options(scipen=999)
rm(list=ls())
```

# Data preparation A: selection of countries and general modifications (as tolower), goes to intermediate data folder

## Libraries

```{r }
library(dplyr)
library(haven)

```

## Get data

```{r results='hide'}

issp99<- read_stata("input/data/origin data/dataissp/issp1999.dta")
issp09<- read_stata("input/data/origin data/dataissp/issp2009.dta")
issp19<- read_stata("input/data/origin data/dataissp/issp2019.dta")

```

## Select Chile
```{r seleccionar Chile}

# 1999

sjmisc::find_var(issp99,"Chile",search = "all")
sjlabelled::get_labels(issp99$v3,values = "as.name")

cl99 = issp99 %>% dplyr::filter(v3==30)

# 2009

sjmisc::find_var(issp09,"Chile",search = "all")
sjlabelled::get_labels(issp09$V4,values = "as.name")

cl09 <- issp09 %>% dplyr::filter(V4==152)

# 2019 ... solo Chile

cl19 <- issp19

```

## Lowercase & save

```{r}
bases=c("cl99","cl09","cl19")

for (i in bases) {
  df.tmp <- get(i)
  names(df.tmp) <- tolower(names(df.tmp)) 
  assign(i, df.tmp)
  save(list=paste(i, sep=""), file = (paste("input/data/intermediate/", i,".RData", sep="")))
}

```