### 2020-01-16 script/data-prep-B.Rmd

---
title: 'Data preparation B'
author: ""
output:
    html_document:
        smart: false
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
```

# Data preparation B: selects & recodes

## Libraries

```{r libraries}
library(sjlabelled)
library(dplyr)
library(haven)
library(descr)
library(sjmisc)
library(stargazer)

```

## Variables:

a- salaries perceived & just worker & manager
b- education
c- income
d- redistribution
e- estatus subjetivo

## Year 1999

### Load data 1999

```{r }
load("input/data/intermediate/cl99.RData") # cl99
```

### a- salaries perceived & just worker & manager 1999

```{r }

# Find vars

find_var(cl99,"worker",search = "all") # v21: perceived worker, v31 just worker unrecoded
find_var(cl99,"chairman",search = "all") # v16: perceived chairman, v26 just chairman unrecoded

# Create object with vars & rename

dat99 <- cl99 %>% select(v21,v31,v16,v26) %>% rename(salperger=v16,
               salperobr=v21,
               saljusger=v26,
               saljusobr=v31)

# Recodes

stargazer(dat99, type="text")

dat99[][dat99  >= 99999999998] <- NA # missings

dat99 <- dat99 %>%
  mutate(salperobr=replace(salperobr, salperobr <= 40000 | salperobr>= 1000000, NA)) %>%
  mutate(salperger=replace(salperger, salperger <= 250000 | salperger>= 100000001, NA)) %>%
  mutate(saljusobr=replace(saljusobr, saljusobr <= 40000, NA)) %>%
  mutate(saljusger=replace(saljusger, saljusger <= 100000, NA))

#---Brecha salarial percibida
dat99$gap_perc   <-  as.numeric(dat99$salperger/dat99$salperobr) # diferencia total
dat99$lngap_perc <-  as.numeric(log(dat99$gap_perc))                 # diferencia log

#---Brecha salarial justa
dat99$gap_just   <-  as.numeric(dat99$saljusger/dat99$saljusobr) # diferencia total
dat99$lngap_just <-  as.numeric(log(dat99$gap_just))                # diferencia log
dim(na.omit(dat99))

summary(dat99)
```

### b - education 1999

```{r}

# Find vars

find_var(cl99,"degr",search = "all") # degree: perceived worker, v31 just worker
frq(as.factor(cl99$x_degr))
frq(as.factor(cl99$degree))

# Problema: no se encuentran los codigos originales en codebook, se derivan de la pregunta father education v317, que tiene las 7 categorías unesco asumiendo que las 7 de degree son las mismas:

# 1. None
# 2. Incomplete primary
# 3. Primary completed
# 4. Incomplete secondary (also includes completed qualifications at a level below academic secondary school
# completion – for example skilled manual qualifications given by trade schools.)
# 5. Secondary completed
# 6. Some tertiary education but less than a university Bachelor’s degree (Includes incomplete university. Also
# includes completed post-secondary courses below university Bachelor’s degree level. Examples are qualifications
# from US junior college, polytechnical institutes in many countries, and other similar post-secondary education.)
# 7. University completed (Bachelor’s degree or higher)

# Codificación original Francisco queda stand-by:

# #dat99 <- mutate(dat99, educ_rec = car::recode(dat99$educ, "3001:3004 = 1;  = 2;3008=2; 3010 = 3;3007 = 4"))
# dat99$educ_rec[dat99$educ_rec== 3001:3004 ] <- 1
# dat99$educ_rec[dat99$educ_rec== 3005:3006 ] <- 2
# dat99$educ_rec[dat99$educ_rec== 3008] <- 2
# dat99$educ_rec[dat99$educ_rec== 3010] <- 3
# dat99$educ_rec[dat99$educ_rec== 3007] <- 4

edcl99 <- cl99 %>% select(degree) %>% mutate(educ_rec=rec(degree,rec = "1:4=1; 5=2 ; 6=3 ; 7=4"))

names(dat99)

dat99<-cbind(dat99,edcl99 %>% select(educ_rec)) # add educ to table

summary(dat99)

dat99 %>% group_by(educ_rec) %>% summarize(m=median(gap_perc, na.rm = TRUE))
```

### c-income 1999

```{r}

find_var(cl99,"income",search = "all") # v21: perceived worker, v31 just worker

freq(cl99$incomer) # deciles family
freq(cl99$rincomer) # 10 categories respondent, more missings
freq(cl99$income) # 14 categories family
freq(cl99$rincome) # categories respondent

# take income, impute middle range, then divide by persons in household, then deciles

cl99$income_c= rec(cl99$income,rec =
      "1=45000;
      2=105500;
      3=135500;
      4=165500;
      5=195500;
      6=225500;
      7=265500;
      8=240500;
      9=495500;
      10=800500;
      11=1250000;
      12=1750000;
      13=2500000;
      14=4000000;
      999998=NA;
      999999=NA")

freq(cl99$income_c) # ok

find_var(cl99,"household",search = "all") # persons in household
freq(cl99$hompop)

dat99$ingreso_pc= cl99$income_c/cl99$hompop

# Generar quintiles de ingreso

dat99 %>% mutate(Quint = ntile(ingreso_pc,5)) -> dat99

# Check
freq( dat99$Quint)

```

### d - redistribution 1999

```{r}
find_var(cl99,"income",search = "all") # v35
freq(cl99$v35)

get_labels(cl99$v35,values = "as.name")

dat99$redis <-rec(cl99$v35, rec="8:9=NA;else=copy")
dat99$redis <-rec(dat99$redis, rec="rev")
freq(dat99$redis)
dat99$redis <-set_labels(dat99$redis,labels= c("Muy en desacuerdo"=1, "En descacuerdo"=2,"Ni acuerdo ni desacuerdo"=3, "De acuerdo"=4, "Muy de acuerdo"=5))
get_labels(dat99$redis, values="as.name")

#dummy redis
dat99$redis_dummy<-ifelse(dat99$redis>3,1,0)
dat99$redis_dummy <-set_labels(dat99$redis_dummy,labels= c("En desacuerdo"=0, "De acuerdo"=1))
freq(dat99$redis_dummy)

freq(to_label(dat99$redis))
```
### e - estatus subjetivo

```{r}
table(cl99$v46)
cl99$v46 <- car::recode(cl99$v46,"c(97,98,99)=NA",as.numeric = TRUE)
cl99$v46 <- sjmisc::rec(cl99$v46,rec="rev")
freq(cl99$v46) # OK esatus subjetivo 1999
dat99$ess <- cl99$v46 # save estatus subjetivo en dat99
```

### Save 99 data
```{r}
save(dat99,file = "input/data/processed/dat99.RData")
names(dat99)
```

## Year 2009

### Load data 2009
```{r }
load("input/data/intermediate/cl09.RData") # cl09
```

### a- salaries perceived & just worker & manager 2009
```{r }

# Find vars

find_var(cl09,"worker",search = "all") # v25: perceived worker, v30 just worker unrecoded
find_var(cl09,"chairman",search = "all") # v23: perceived chairman, v28 just chairman unrecoded

# Create object with vars & rename

dat09 <- cl09 %>% select(v25,v30,v23,v28) %>% rename(salperger=v23,
               salperobr=v25,
               saljusger=v28,
               saljusobr=v30)

# Recodes
stargazer(dat09, type="text")

dat09[][dat09  >= 99999999998] <- NA # missings

dat09 <- dat09 %>%
  mutate(salperobr=replace(salperobr, salperobr <= 40000 | salperobr>= 1000000, NA)) %>%
  mutate(salperger=replace(salperger, salperger <= 250000 | salperger>= 100000001, NA)) %>%
  mutate(saljusobr=replace(saljusobr, saljusobr <= 40000, NA)) %>%
  mutate(saljusger=replace(saljusger, saljusger <= 100000, NA))

#---Brecha salarial percibida
dat09$gap_perc   <-  as.numeric(dat09$salperger/dat09$salperobr) # diferencia total
dat09$lngap_perc <-  as.numeric(log(dat09$gap_perc))                 # diferencia log

#---Brecha salarial justa
dat09$gap_just   <-  as.numeric(dat09$saljusger/dat09$saljusobr) # diferencia total
dat09$lngap_just <-  as.numeric(log(dat09$gap_just))                # diferencia log
dim(na.omit(dat09))

stargazer(dat09, type="text")
```

### b - education 2009
```{r}

# Find vars

##sjPlot::view_df(cl09, show.frq = T)

find_var(cl09,"degr",search = "all") # cl_degr:
frq(as.factor(cl09$cl_degr))

# Problema: no se encuentran los codigos originales en codebook, se derivan de la pregunta father education v317, que tiene las 7 categorías unesco asumiendo que las 7 de degree son las mismas:

# 1. None
# 2. Incomplete primary
# 3. Primary completed
# 4. Incomplete secondary (also includes completed qualifications at a level below academic secondary school
# completion – for example skilled manual qualifications given by trade schools.)
# 5. Secondary completed
# 6. Some tertiary education but less than a university Bachelor’s degree (Includes incomplete university. Also
# includes completed post-secondary courses below university Bachelor’s degree level. Examples are qualifications
# from US junior college, polytechnical institutes in many countries, and other similar post-secondary education.)
# 7. University completed (Bachelor’s degree or higher)

# Codificación original Francisco queda stand-by:

# #dat99 <- mutate(dat99, educ_rec = car::recode(dat99$educ, "3001:3004 = 1;  = 2;3008=2; 3010 = 3;3007 = 4"))
# dat99$educ_rec[dat99$educ_rec== 3001:3004 ] <- 1
# dat99$educ_rec[dat99$educ_rec== 3005:3006 ] <- 2
# dat99$educ_rec[dat99$educ_rec== 3008] <- 2
# dat99$educ_rec[dat99$educ_rec== 3010] <- 3
# dat99$educ_rec[dat99$educ_rec== 3007] <- 4

edcl09 <- cl09 %>% select(cl_degr) %>% mutate(educ_rec=rec(cl_degr,rec = "1:4=1; 5:6=2; 8=2 ; 9=3 ; 7=4"))

names(edcl09)

dat09<-cbind(dat09,edcl09 %>% select(educ_rec)) # add educ to table

summary(dat09)

dat09 %>% group_by(educ_rec) %>% summarize(m=median(gap_perc, na.rm = TRUE))
```

### c-income 2009

```{r}

find_var(cl09,"income",search = "all") # cl_inc

freq(cl09$cl_inc) # deciles family

# take income, impute middle range, then divide by persons in household, then deciles

cl09$income_c= rec(cl09$cl_inc,rec =
      "20000=45000;
      45000=105500;
      66000=135500;
      90000=165500;
      122000=195500;
      155000=225500;
      200000=265500;
      250000=240500;
      310000=495500;
      400000=800500;
      750000=1250000;
      1500000=1750000;
      2500000=2500000;
      3100000=4000000;
      9999998=NA;
      9999999=NA")

freq(cl99$income_c) # ok

find_var(cl09,"household",search = "all") # persons in household
freq(cl09$hompop)

dat09$ingreso_pc= cl09$income_c/cl09$hompop

# Generar quintiles de ingreso

dat09 %>% mutate(Quint = ntile(ingreso_pc,5)) -> dat09

# Check
freq( dat09$Quint)
```

### d - redistribution 2009

```{r}
find_var(cl09,"income",search = "all") # v33
freq(cl09$v33)
get_labels(cl09$v33,values = "as.name")

dat09$redis <-rec(cl09$v33, rec="8:9=NA;else=copy")
dat09$redis <-rec(dat09$redis, rec="rev")
freq(dat09$redis)
dat09$redis <-set_labels(dat09$redis,labels= c("Muy en desacuerdo"=1, "En descacuerdo"=2,"Ni acuerdo ni desacuerdo"=3, "De acuerdo"=4, "Muy de acuerdo"=5))
get_labels(dat09$redis, values="as.name")

#dummy redis
dat09$redis_dummy<-ifelse(dat09$redis>3,1,0)
dat09$redis_dummy <-set_labels(dat09$redis_dummy,labels= c("En desacuerdo"=0, "De acuerdo"=1))
freq(dat09$redis_dummy)
```

### e - estatus subjetivo

```{r}
table(cl09$v44)
cl09$v44 <- car::recode(cl09$v44,"c(98,99)=NA",as.numeric = TRUE)
freq(cl09$v44) # OK esatus subjetivo 2009
dat09$ess <- cl09$v44 # save estatus subjetivo en dat09
freq(dat09$ess)
```

### Save 09 data
```{r}
save(dat09,file = "input/data/processed/dat09.RData")
```

## Year 2019

### Load data 2019

```{r }
load("input/data/intermediate/cl19.RData") # cl19
```

### a- salaries perceived & just worker & manager 2019

```{r }

# Find vars

find_var(cl19,"obrero",search = "all") # m2_p2_obrero: perceived worker, m2_p3_obrero: just worker
find_var(cl19,"presidente",search = "all") # m2_p2_presidente: perceived chairman, m2_p3_presidente: just chairman

# Create object with vars & rename

dat19 <- cl19 %>% select(m2_p2_presidente,m2_p2_obrero,m2_p3_obrero,m2_p3_presidente) %>% rename(salperger=m2_p2_presidente,
               salperobr=m2_p2_obrero,
               saljusger=m2_p3_presidente,
               saljusobr=m2_p3_obrero)

# Recodes

stargazer(dat19, type="text")

dat19[][dat19  >= 99999999998] <- NA # missings

dat19 <- dat19 %>%
  mutate(salperobr=replace(salperobr, salperobr <= 40000 | salperobr>= 1000000, NA)) %>%
  mutate(salperger=replace(salperger, salperger <= 250000 | salperger>= 100000001, NA)) %>%
  mutate(saljusobr=replace(saljusobr, saljusobr <= 40000, NA)) %>%
  mutate(saljusger=replace(saljusger, saljusger <= 100000, NA))

#---Brecha salarial percibida
dat19$gap_perc   <-  as.numeric(dat19$salperger/dat19$salperobr) # diferencia total
dat19$lngap_perc <-  as.numeric(log(dat19$gap_perc))                 # diferencia log

#---Brecha salarial justa
dat19$gap_just   <-  as.numeric(dat19$saljusger/dat19$saljusobr) # diferencia total
dat19$lngap_just <-  as.numeric(log(dat19$gap_just))                # diferencia log
dim(na.omit(dat19))

hist(dat19$saljusger)
hist(dat19$saljusobr)
hist(dat19$gap_just)
hist(dat19$gap_perc)

```

### b - education 2019

```{r}
# Find vars

find_var(cl19,"Educ",search = "all") # ds_p4:
frq(as.factor(cl19$ds_p4))

# Problema: no se encuentran los codigos originales en codebook, se derivan de la pregunta father education v317, que tiene las 7 categorías unesco asumiendo que las 7 de degree son las mismas:

# 1. None
# 2. Incomplete primary
# 3. Primary completed
# 4. Incomplete secondary (also includes completed qualifications at a level below academic secondary school
# completion – for example skilled manual qualifications given by trade schools.)
# 5. Secondary completed
# 6. Some tertiary education but less than a university Bachelor’s degree (Includes incomplete university. Also
# includes completed post-secondary courses below university Bachelor’s degree level. Examples are qualifications
# from US junior college, polytechnical institutes in many countries, and other similar post-secondary education.)
# 7. University completed (Bachelor’s degree or higher)

# Codificación original Francisco queda stand-by:

# #dat99 <- mutate(dat99, educ_rec = car::recode(dat99$educ, "3001:3004 = 1;  = 2;3008=2; 3010 = 3;3007 = 4"))
# dat99$educ_rec[dat99$educ_rec== 3001:3004 ] <- 1
# dat99$educ_rec[dat99$educ_rec== 3005:3006 ] <- 2
# dat99$educ_rec[dat99$educ_rec== 3008] <- 2
# dat99$educ_rec[dat99$educ_rec== 3010] <- 3
# dat99$educ_rec[dat99$educ_rec== 3007] <- 4

edcl19 <- cl19 %>% select(ds_p4) %>% mutate(educ_rec=rec(ds_p4,rec = "0:3=1; 4:5=2 ;7=2 ; 6=3 ; 8:9=4"))

dat19<-cbind(dat19,edcl19 %>% select(educ_rec)) # add educ to table

summary(dat19)
table(dat19$educ_rec)

dat19 %>% group_by(educ_rec) %>% summarize(m=median(gap_perc, na.rm = TRUE))
```

### c-income 2019

```{r}

find_var(cl19,"este hogar",search = "all") # ds_p39: perceived worker, v31 just worker

freq(cl19$ds_p39) # deciles family

# take income, impute middle range, then divide by persons in household, then deciles

cl19$income_c= rec(cl19$ds_p39,rec =
      "1=45000;
      2=105500;
      3=135500;
      4=165500;
      5=195500;
      6=225500;
      7=265500;
      8=240500;
      9=495500;
      10=800500;
      11=1250000;
      12=1750000;
      13=2500000;
      14=4000000;
      88=NA;
      99=NA")

freq(cl19$income_c) # ok

find_var(cl19,"viven",search = "all") # ds_p34
freq(cl19$ds_p34)

dat19$ingreso_pc= cl19$income_c/cl19$ds_p34

# Generar quintiles de ingreso

dat19 %>% mutate(Quint = ntile(ingreso_pc,5)) -> dat19

# Check
freq( dat19$Quint)
```

### d - redistribution 2019

```{r}
find_var(cl19,"responsabilidad",search = "all") # m2_p4_2
freq(cl19$m2_p4_2 )
get_labels(cl19$m2_p4_2 ,values = "as.name")

dat19$redis <-rec(cl19$m2_p4_2, rec="8:9=NA;else=copy")
dat19$redis <-rec(dat19$redis, rec="rev")
freq(dat99$redis)
dat19$redis <-set_labels(dat19$redis,labels= c("Muy en desacuerdo"=1, "En descacuerdo"=2,"Ni acuerdo ni desacuerdo"=3, "De acuerdo"=4, "Muy de acuerdo"=5))
get_labels(dat19$redis, values="as.name")

#dummy redis
dat19$redis_dummy<-ifelse(dat19$redis>3,1,0)
dat19$redis_dummy <-set_labels(dat19$redis_dummy,labels= c("En desacuerdo"=0, "De acuerdo"=1))
freq(dat99$redis_dummy)

freq(to_label(dat99$redis))
```

### e - estatus subjetivo

```{r}
table(cl19$m2_p13a)
cl19$m2_p13a <- car::recode(cl19$m2_p13a,"c(88,99)=NA",as.numeric = TRUE)
freq(cl19$m2_p13a) # OK esatus subjetivo 2019
dat19$ess <- cl19$m2_p13a # save estatus subjetivo en dat19
freq(dat19$ess)
```

### Save 19 data
```{r}

save(dat19,file = "input/data/processed/dat19.RData")
```

### Save all years data
```{r}
load("input/data/processed/dat99.RData")
load("input/data/processed/dat09.RData")
load("input/data/processed/dat19.RData")

dat99$year <- 1999
dat09$year <- 2009
dat19$year <- 2019
clw3<- bind_rows(dat99,dat09)
clw3<- bind_rows(clw3,dat19)
save(clw3,file = "input/data/processed/clw3.Rdata")
names(clw3)
```

<!-- # tratamiento CASEN 2000-2017 -->

<!-- ##Casen 2000 -->
<!-- ```{r} -->
<!-- load(file="input/data/original/datacasen/casen2000.RData") -->

<!-- ##sjPlot::view_df(select(cs2000, "o8"), show.values = T, show.labels = T, max.len = 1000) -->
<!--                 #hacer codebook para o8 -->

<!-- #peones de... -->
<!-- casen_obreros2000 = cs2000 %>% dplyr::filter(o8>=9210, o8<=9322) -->
<!-- casen_obreros2000 <- casen_obreros2000 %>% dplyr::select(yopraj) -->
<!-- casen_obreros2000<- na.omit(casen_obreros2000) -->

<!-- #gerentes de -->
<!-- casen_gerente2000 = cs2000 %>% dplyr::filter(o8==1210) #gerente empresa -->
<!-- #casen_gerente2000 <- casen_gerente2000 %>% dplyr::select(yopraj,o13,o10,o28, o11) -->

<!-- casen_gerente2000 = casen_gerente2000 %>% dplyr::filter(o10==1) -->
<!-- casen_gerente2000 = casen_gerente2000 %>% dplyr::filter(o13=="E" ) -->

<!-- brecha_real2000<-(median(casen_gerente2000$yopraj))/(median(casen_obreros2000$yopraj)) -->

<!-- summary(casen_obreros2000$yopraj) -->
<!-- summary(casen_gerente2000$yopraj) -->
<!-- summary(brecha_real2000) -->
<!-- table(casen_gerente2000$o13) -->

<!-- #sjPlot::view_df(casen_gerente2000, show.frq = T, max.len = 100000)#hacer codebook -->
<!-- ``` -->
<!-- ##Casen 2009 -->

<!-- ```{r} -->
<!-- load(file="input/data/original/datacasen/casen2009.RData") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ##sjPlot::view_df(select(cs2009, "c_o12"), show.values = T, show.labels = T, max.len = 1000) -->

<!-- #ocupacion:c_o12 -->
<!-- #ingreso: yopraj -->

<!-- #peones de... -->
<!-- casen_obreros2009 = cs2009 %>% dplyr::filter(c_o12>=9200, c_o12<=9322) -->
<!-- casen_obreros2009 <- casen_obreros2009 %>% dplyr::select(yopraj) -->
<!-- casen_obreros2009<- na.omit(casen_obreros2009) -->

<!-- #gerentes de -->
<!-- casen_gerente2009 = cs2009 %>% dplyr::filter(c_o12==1210) #gerente general empresa -->
<!-- casen_gerente2009 <- casen_gerente2009 %>% dplyr::select(yopraj,o14,o23) -->
<!-- casen_gerente2009 = casen_gerente2009 %>% dplyr::filter(o14=="F")    #gran empresa | cuando se utilizan los dos filtros se llega  a una brecha de 52 -->
<!-- casen_gerente2009 = casen_gerente2009 %>% dplyr::filter(o23==1)      # filtro por ser patron en la empresa. -->
<!-- casen_gerente2009<-na.omit(casen_gerente2009) -->
<!-- table(casen_gerente2009$o14) -->
<!-- #sjPlot::view_df(casen_gerente2009, show.frq = T, max.len = 1000) -->

<!-- brecha_real2009<-(median(casen_gerente2009$yopraj))/(median(casen_obreros2009$yopraj)) -->

<!-- summary(casen_obreros2009$yopraj) -->
<!-- summary(casen_gerente2009$yopraj) -->
<!-- brecha_real2009 -->
<!-- #sjPlot::view_df(casen_gerente2009, show.frq = T, max.len = 300) # hacer codebook -->
<!-- ``` -->

<!-- ```{r} -->
<!-- load(file="input/data/original/datacasen/casen2017.RData") -->
<!-- ``` -->

<!-- ## Casen 2017 -->
<!-- ```{r} -->
<!-- #codecasen17<-#sjPlot::view_df(cs2017)#hacer codebook -->
<!-- ##sjPlot::view_df(cs2017, show.values = T, show.labels = T, max.len = 1000) -->
<!--                 #hacer codebook para o8 -->

<!-- #peones de... -->
<!-- casen_obreros2017 = cs2017 %>% dplyr::filter(oficio4>=9211, oficio4<=9322) -->
<!-- casen_obreros2017 <- casen_obreros2017 %>% dplyr::select(y0101, o23) -->
<!-- casen_obreros2017<- na.omit(casen_obreros2017) -->

<!-- #gerentes de -->
<!-- #casen_gerente2017 = cs2017 %>% dplyr::filter(oficio4 %in% c(1210:1319)) # Filtro con todos los gerentes -->
<!-- casen_gerente2017 = cs2017 %>% dplyr::filter(oficio4==1210) #gerente empresa -->
<!-- casen_gerente2017 <- casen_gerente2017 %>% dplyr::select(oficio4, y0101,o23, o15,yrut,y0701,y0801) -->
<!-- casen_gerente2017 = casen_gerente2017 %>% dplyr::filter(o23=="F") #gran empresa > 100 trabajadores -->
<!-- casen_gerente2017 = casen_gerente2017 %>% dplyr::filter(o15 == 1) # empleador -->

<!-- #casen_gerente2017 = casen_gerente2017 %>% dplyr::filter(y0701 <= 1000000) # empleador -->

<!-- # yrut # retiro de utilidades -->
<!-- #y0701 #independiente principal -->
<!-- #table(casen_gerente2017$o15) -->
<!-- # #sjPlot::view_df(casen_gerente2017, show.frq = T, max.len = 1000) -->

<!-- brecha_real2017<-(median(casen_gerente2017$y0701,na.rm = T))/(median(casen_obreros2017$y0101,na.rm = T)) -->

<!-- summary(casen_obreros2017$y0101) -->
<!-- table(casen_gerente2017$y0701) -->
<!-- brecha_real2017 -->

<!-- summary(cs2017$y0101) -->
<!-- summary(cs2017$y0701) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- brecha_salarial<- rbind(psych::describe(brecha_real2000), -->
<!--                psych::describe(brecha_real2009), -->
<!--                psych::describe(brecha_real2017)) -->

<!-- save(brecha_salarial, file = "input/data/processed/dataproccasen/brecha.Rdata") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- salario_gerente<- rbind(psych::describe(casen_gerente2000$yopraj), -->
<!--                psych::describe(casen_gerente2009$yopraj), -->
<!--                psych::describe(casen_gerente2017$y0701)) -->

<!-- save(salario_gerente, file = "input/data/processed/dataproccasen/salario_gerente.Rdata") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- salario_obrero<- rbind(psych::describe(casen_obreros2000$yopraj), -->
<!--                psych::describe(casen_obreros2009$yopraj), -->
<!--                psych::describe(casen_obreros2017$y0101)) -->

<!-- save(salario_obrero, file = "input/data/processed/dataproccasen/salario_obrero.Rdata") -->
<!-- ``` -->

<!-- ``` -->